---
title: "Explore data using SQL"
subtitle: "A two-day workshop at SNEHA, Mumbai"
institute: "CivicDataLab"
date: "14th - 15th November, 2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo("https://drive.google.com/uc?id=1WbdosVvi6SBjYtEF6kHj-f0wotdHMRgw",link_url = "https://civicdatalab.in",exclude_class = NULL)
```

```{r xaringan-fit-screen, echo=FALSE}
xaringanExtra::use_fit_screen()
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```

## Workshop Overview

Topics to be covered:




Outline:

Day 1
1. Overview - Agenda for the two-day workshop
2. Working with structured datasets - Data Representation / Data Types
3. Data Types - Examples and Use-cases
4. Spreadsheets vs Databases
5. Tools / Installation - About Postgres
6. Reading and Writing SQL (Select, Sorting and Removing Duplicates, Filtering, Aggregation)
7. SQL - Basic Examples
8. A SQL project - Best practices

Day 2
1. Day 1 - Recap
2. SQL - Examples - Contd.
3. Writing SQL (Calculating New Values, Missing Data, ) 
4. SQL - Case Study (Using datasets managed by SNEHA)
5. SQL - Advance (Working with Dates, Combining Datasets, Joins, Case When, Subqueries)
6. SQL puzzle - https://mystery.knightlab.com/
7. Resources to learn and practice SQL

---

class: center, middle

# Queries and Feedback
