---
title: "Module 2 - Session 3 - Data exploration"
subtitle: "Working effectively with data"
institute: "CivicDataLab"
date: "2021/08/24 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo("https://drive.google.com/uc?id=1WbdosVvi6SBjYtEF6kHj-f0wotdHMRgw",link_url = "https://civicdatalab.in",exclude_class = NULL)
```

```{r xaringan-fit-screen, echo=FALSE}
xaringanExtra::use_fit_screen()
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(fontawesome)
style_mono_accent(base_color = "#A63A50",colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF",
  yellow = "#F1A208"
  ),link_decoration="underline",table_row_even_background_color = "#FFFFFF")

```


# Exercise - Exploring data from eCourts

**Dataset** - [Link](https://www.dropbox.com/sh/hkcde3z2l1h9mq1/AADRe-BuBQ92ozAJiG7YERdCa?dl=0) - _The database contains 81.2 million cases_

**Source**: [Devdatalab](http://www.devdatalab.org/judicial-data)

**Objective**:

- Understand how the data is structured
- Import the data in a database
- Explore the sample datasets
- Find out the total cases present for each district for the year 2018

**Tags**

.bg-yellow[.black[_database_]]  .bg-yellow[.black[_large-datasets_]]  .bg-yellow[.black[_sqlite_]]  .bg-yellow[.black[_eCourts_]]  

---

# Exercise - Using Databases

- Install SQLite DB Browser
- Create a new database
- Load the judges_clean dataset in the DB
- Find the distribution of male/female judges in **Bengaluru** district court where judge position is _chief metropolitan magistrate_
- Save the file, as CSV, in the drive

---

# SQL - Recap 👇

<iframe src="https://https://civicdatalab.in/Working-with-Data-Workshops/modules/module_2_data_exploration/session-2.html" frameborder="0" height="480" width="640">
</iframe>



