---
title: "Module 2 - Session 1 - Data exploration"
subtitle: "Working effectively with data"
institute: "CivicDataLab"
date: "2021/08/03 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo("https://drive.google.com/uc?id=1WbdosVvi6SBjYtEF6kHj-f0wotdHMRgw",link_url = "https://civicdatalab.in",exclude_class = NULL)
```

```{r xaringan-fit-screen, echo=FALSE}
xaringanExtra::use_fit_screen()
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(fontawesome)
style_mono_accent(base_color = "#A63A50",colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF",
  yellow = "#F1A208"
  ),link_decoration="underline",table_row_even_background_color = "#FFFFFF")

```

# Past sessions

| Session | Topic | Date | Slide Deck | Session Recording | Notes |
| :-----: | :---: | :--: | :--------: | :---------------: | :---: |
| 1 | Data Collection | 16/07/21 | [`r fa(name = "firefox-browser", fill = "steelblue", height = "20px")`](https://civicdatalab.in/Working-with-Data-Workshops/modules/module_1_data_collection/session-1.html) [`r fa(name = "file-pdf", fill = "steelblue", height = "20px")`](https://civicdatalab.in/Working-with-Data-Workshops/modules/module_1_data_collection/session-1.pdf) | [`r fa(name = "youtube", fill = "steelblue", height = "20px")`](https://youtu.be/DwfiBPyxlLM) | [`r fa(name = "file-word", fill = "steelblue", height = "20px")`](https://docs.google.com/document/d/1ZBMfqCvjPmqRQr2cUpVazK81SYK4GFJK3qf3uXxJoOY/edit?usp=sharing) |
| 2 | Web scraping & Data Standards | 26/07/21 | [`r fa(name = "firefox-browser", fill = "steelblue", height = "20px")`](https://civicdatalab.in/Working-with-Data-Workshops/modules/module_1_data_collection/session-2/session-2.html) [`r fa(name = "file-pdf", fill = "steelblue", height = "20px")`](https://civicdatalab.in/Working-with-Data-Workshops/modules/module_1_data_collection/session-2/session-2.pdf) | [`r fa(name = "youtube", fill = "steelblue", height = "20px")`](https://www.youtube.com/watch?v=IaA_qClEFMg) | [`r fa(name = "file-word", fill = "steelblue", height = "20px")`](https://docs.google.com/document/d/1O5AbZhz3rAv9Nyy1BF4GWuvS8-AfR8bzbzU4RzwxTRw/edit) |

.bg-washed-red.b--dark-red.ba.bw2.br3.shadow-5.ph4.mt5[
Thanks for sharing your feedback and suggestions with us. If you haven't got a chance to do it yet, the form is still open here -> [https://forms.gle/6pzx8XKBZ7YUGV336](https://forms.gle/6pzx8XKBZ7YUGV336) 

]

---
class: center, middle

# Data exploration Use-Cases

---

# Exploring state level mortality data

**Dataset** - [Link](https://www.dropbox.com/sh/y949ncp39towulf/AABs8_dECTzr38GdS7BneTH7a?dl=0)

**Source**: [Devdatalab](http://www.devdatalab.org/covid)

**Objective**:

- Import data in google sheets
- Compare mortality data across states from 2019 - 2021
- Compare mortality data for Karnataka across years, as shown [here](https://devdatalab.medium.com/making-sense-of-excess-mortality-f26cffe3643e)

**Tags**

.bg-yellow[.black[_google-sheets_]]  .bg-yellow[.black[_data-viz_]]  .bg-yellow[.black[_pivot_]]  .bg-yellow[.black[_data-structure_]]

---

# Most cited IPC acts/sections across courts

**Dataset** - [Link](https://justicehub.in/dataset/e48994b7-e77a-4fa1-869f-923dec3e4636/resource/02005017-cd9d-4fec-94d4-243b5e4f9fcb/download/top_sections.csv)

**Source**: [Contributed by IndianKanoon on the Justice Hub](https://justicehub.in/dataset/indian-kanoon-statistics)

**Objective**:

- Import data in google sheets
- Filter data for IPC sections
- Find how many sections contribute to 50% of the cases
- Add a column for the section title

**Tags**

.bg-yellow[.black[_google-sheets_]]  .bg-yellow[.black[_functions_]]  .bg-yellow[.black[_lookup_]]  .bg-yellow[.black[_scraping_]]  .bg-yellow[.black[_XML_]]  .bg-yellow[.black[_JSoN_]]


---

# Working with databases

- Dealing with large datasets


- Platform agnostic


- Programming language agnostic


- Easy to share and maintain as compared to multiple data files


---

# Exploring data from eCourts

**Dataset** - [Link](https://www.dropbox.com/sh/hkcde3z2l1h9mq1/AADRe-BuBQ92ozAJiG7YERdCa?dl=0) - _The database contains 81.2 million cases_

**Source**: [Devdatalab](http://www.devdatalab.org/judicial-data)

**Objective**:

- Understand how the data is structured
- Import the data in a database
- Explore the sample datasets
- Find out the total cases present for each district for the year 2018

**Tags**

.bg-yellow[.black[_database_]]  .bg-yellow[.black[_large-datasets_]]  .bg-yellow[.black[_sqlite_]]  .bg-yellow[.black[_eCourts_]]  

---

# Tools 

1. [Sublime Text](https://www.sublimetext.com/) - _For text processing_
2. [SQLite Browser](https://sqlitebrowser.org/) - _For working with database_
3. [JSon Editor](https://jsoneditoronline.org/) - _For editing/viewing JSoN files_
4. [XML to JSON](https://codebeautify.org/xmltojson) - _For converting XML to JSON_
5. [jq](https://stedolan.github.io/jq/) - _For working with JSON files_
6. [Agenty](https://chrome.google.com/webstore/detail/agenty-advanced-web-scrap/gpolcofcjjiooogejfbaamdgmgfehgff) - _Chrome extension for scraping data_
7. [CSVLint](https://csvlint.io/) - _For validating CSV files_

---

# Do try this at home

.pull-left[

**Exercise - 1**

- [Link](https://nalsa.gov.in/dashboard/) to NALSA dashboard
- Create a CSV file with variables available under the Victim Compensation Schemes table for these states: 
    - Delhi
    - Maharashtra
    - Karnataka
    - West Bengal
    - Uttar Pradesh
- Create a chart to compare the yearly compensation numbers between these states
- Create a folder [here](https://drive.google.com/drive/u/0/folders/1W1t0j1NETZDQ5F7j7Ts6tDNiMQUTKGXD) and upload the dataset (including the chart) 

]

.pull-right[

**Exercise - 2**

- Install SQLite DB Browser
- Create a new database
- Load the judges_clean dataset in the DB
- Find the distribution of male/female judges in **Bengaluru** district court where judge position is _chief metropolitan magistrate_
- Save the file, as CSV, in the drive

]

---

# Other Resources

- [Working with SQLite](https://github.com/sqlitebrowser/sqlitebrowser/wiki)
- [SQL Tutorial](https://www.w3schools.com/sql/)
- [Databases and SQL practice](https://swcarpentry.github.io/sql-novice-survey/)