---
title: "Exploring datasets - Examples"
subtitle: "Workshop at NLU Odisha"
institute: "CivicDataLab"
date: "2022/08/08 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo("https://drive.google.com/uc?id=1WbdosVvi6SBjYtEF6kHj-f0wotdHMRgw",link_url = "https://civicdatalab.in",exclude_class = NULL)
```

```{r xaringan-fit-screen, echo=FALSE}
xaringanExtra::use_fit_screen()
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```

# Crime in India

Source: [Wikipedia](https://en.wikipedia.org/wiki/Crime_in_India)

## Objective:

1. Download the table _Absolute number of reported crimes & crime rates across India_
2. Calculate _total reported crimes from 2016-2019_ in all states
3. Calculate _percent share of each state_ in reported crimes
4. Arrange the results in decreasing order of percent share

---

# Most cited acts and sections

Source: [Justice Hub - Indian Kanoon Statistics](https://justicehub.in/dataset/indian-kanoon-statistics)


## Objective:

1. Import data in google sheets
2. Filter data for IPC sections
3. Find how many sections contribute to 50% of the cases
4. Add a column for the section title

---

# Mortality data from DevDataLab

Source: Dropbox

Link: https://www.dropbox.com/sh/y949ncp39towulf/AACJRhJs6uzbjkB6uM5O9mKga/mortality/csv?dl=0&subfolder_nav_tracking=1

## Objective:

1. Import data in google sheets
2. Compare mortality data across states from 2019 - 2021
3. Compare mortality data for Karnataka across years, as shown [here](https://devdatalab.medium.com/making-sense-of-excess-mortality-f26cffe3643e)

---

# Judges of the Orissa High Court

Source: Justice Hub

Link: https://justicehub.in/dataset/khoj-orissa/resource/92c89ebb-3612-4ad5-a327-d55f65327577

## Objective:

1. Import data in google sheets
2. Find the number of Male and Female judges
3. Find the percentage of judges where _Parent High court_ is Odisha High Court
4. Find the total number of judges appointed each year
5. Find the total number of judges retiring each year

---

# India at CommonWealth Games 

Source: Wikipedia

Link: https://en.wikipedia.org/wiki/India_at_the_Commonwealth_Games

## Objective:

1. Find the table titled - _Overall Medal Tally By Sport_
2. Import data in Microsoft Excel
3. Find the percentage share of total medals by sports
4. Arrange the sports in decreasing order of percentage share of total medals
5. Find the percentage share of gold, silver and bronze medals by each sport
6. Remove Grand Totals from Columns
7. Highlight the cells with the largest contribution to gold, silver and bronze medals respectively

---

class: center, middle

# Queries and Feedback
